\bf Range Analysis for the IonMonkey JavaScript engine

### 2012-05-02 project-writings/paper.tex
cite the benchmarks and stats

There are three benchmark suites which are generally used to measure JavaScript
JIT compilers: the v8bench suite developed by Google, the SunSpider benchmark
suite developed by Apple, and the Kraken benchmark suite developed by Mozilla.

We ran each benchmark suite on IonMonkey, both with and without our range
analysis pass, and measured the performance differences. The results are
summarized in Figures \ref{fig:v8bench}, \ref{fig:sunspider}, and \ref{fig:kraken}.

@@

Further optimizations are possible using our range analysis framework. If the
range associated with an instruction becomes empty, then the block has been
proven dead and can be eliminated. Control flow operations can be proven
redundant as well, further eliminating computation and dead basic blocks. These
graph manipulations are currently not available in the IonMonkey compiler,
which is in active development but not yet complete.

Symbolic range propagation, where the bounds are expressed in terms of
definitions rather than integer constants, is left as future work. Symbolic
range propagation would allow more sophisticated removal of bounds checks,
which are generally not based on constants.

In summary, we augment the IonMonkey compiler with range information and begin
to use this information for the purposes of optimization. We leave more
sophisticated uses of the analysis as future work.

>>>>

There are three benchmark suites which are generally used to measure JavaScript
JIT compilers: the v8bench suite developed by Google\cite{v8bench}, the SunSpider benchmark
suite developed by Apple\cite{sunspider}, and the Kraken benchmark suite developed by Mozilla\cite{kraken}

We ran each benchmark suite on IonMonkey, both with and without our range
analysis pass, and measured the performance differences. The results are
summarized in Figures \ref{fig:v8bench}, \ref{fig:sunspider}, and
\ref{fig:kraken}. These figures were gathered using the SunSpider test harness,
which gives both a mean and a 95\% confidence interval\cite{sunspider}. Each
test was run 30 times.

@@

On a suite of standard Javascript benchmarks, our range analysis
succeeds in eliminating a large fraction of overflow checks for
integer additions. This translates into a minor overall performance
win on the Kraken benchmark suite and a very minor overall performance
loss on the Sunspider suite. Since we should only increase the
quality of the generated code, any performance losses should be from
increased compilation overhead (Sunspider tests in particular are
quite short). The performance improvements are decent, but not
stellar; we believe this is because Javascript code is generally not
integer-arithmetic-heavy tight loops and because the overflow test
branches are generally predicted correctly.

A number of further optimizations are possible using our range
analysis framework. If the range associated with an instruction
becomes empty, then the block has been proven dead and can be
eliminated. Control flow operations can be proven redundant as well,
further eliminating computation and dead basic blocks. These graph
manipulations are currently not available in the IonMonkey compiler,
which is in active development but not yet complete.

Symbolic range propagation, where the bounds are expressed in terms of
definitions rather than integer constants, is left as future
work. Symbolic range propagation would allow more sophisticated
removal of bounds checks, which are generally not based on
constants. It would also allow for much more potential for eliminating
redundant branches than just constant ranges would.

We believe that we have implemented a worthwhile optimization and a
useful analysis pass that will serve as a good building block for
future IonMonkey optimizations. Our range analysis patches are likely
to land in IonMonkey in the next month, and it is likely that some of
the future work we discuss will be taken up by the IonMonkey team.
